Globally-centered autocovariances in MCMC

### 2020-09-02 Paper/main.tex
Wednesday morning prelim edits. 31 pages.

By \cite[Proposition 1]{song1995optimal},
\[
\Var(\bar{X}) = \dfrac{\Sigma}{n} + \dfrac{\Phi}{n^2} + o(n^{-2})\,.
\]
As a consequence, we get the result 
\[
 \E \left[\hat{\Gamma}_{G,s}(k) \right] =  \left(1- \dfrac{\abs{k}}{n}\right)\left( \Gamma(k) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2} \right) + o(n^{-1})\,.
\]
% As a consequence, if $\Var(\bar{X})$ is finite, then $\Var(\bar{X}) \to 0 \textrm{ as } n \to \infty$. Expectation of $\hat{\Gamma}_{G,s}(k)$ can then broken down as following,
Note that we can also break down $ \E \left[\hat{\Gamma}_{G,s}(k) \right]$ as
 \begin{equation} \label{eq:G-ACF_expec_breakdown}
     \mathbb{E}\left[\hat{\Gamma}_{G,s}(k)\right] = \left(1- \dfrac{\abs{k}}{n}\right)\Gamma(k) + O_1 + O_2\,.
 \end{equation}
%
where,
\begin{align*}
    O_1 &= -\dfrac{\abs{k}}{n}\left[\left(1-\dfrac{1}{m}\right)\left(\dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h)^T + \dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h)\right) - \left(2-\dfrac{1}{m}\right) \left(\dfrac{\Sigma}{n} + \dfrac{\Phi}{n^2}\right)\right] + o(n^{-2})\, ,\\
    O_2 &= -\dfrac{1}{m}\left(\dfrac{\Sigma}{n} + \dfrac{\Phi}{n^2}\right) + o(n^{-2})
\end{align*}
%
Here $O_1 = (\abs{k}/n)\mathcal{O}(1/n) \to 0 \textrm{ and } O_2 = {\mathcal{O}}(1/n) \to 0$ as $n \to \infty$. For a diagonal element of $\Gamma$,
% We observe that this expectation result is very similar to equation \ref{eq:priestly} from \cite{priestley1981spectral} for the asymptotically unbiased estimator of autocovariance except for an additional positive bias term. This helps in countering the inherent negative bias in $\hat{\Gamma}(k)$ and give more informative estimates of autocovariance.  Although, bias term for a fixed $k$ is $o(1)$, it increases with $k$. We have explicitly written the end effect term arising from using the estimator $\bar{\bar{X}}$ for $\mu$. These are small order terms that vanish as $n \to \infty$.
%
\begin{align*}
& \E \left[ \hat{\Gamma}_{G,s}^{ii}\right]\\
& = \mathbb{E}\left[\hat{\Gamma}^{ii}_{s}(k)\right] - \dfrac{\abs{k}}{n}\left[\left(1-\dfrac{1}{m}\right)\left(\dfrac{1}{n}\sum_{h=0}^{n-1} \left[\Gamma^{ii}(h)^T + \Gamma^{ii}(h) \right] \right) - \left(2-\dfrac{1}{m}\right) \Var(\bar{X}_1)^{ii}\right] - \dfrac{\Var(\bar{X}_1)^{ii}}{m}\,.
\end{align*}
Note that $ \hat{\Gamma}_{G,s}(k)$ is less negatively biased as compared to $\hat{\Gamma}_s(k)$. In the presence of positive correlation, the leftover $o(n^{-1})$ term is positive which further counters the negative bias of $\hat{\Gamma}_{G,s}(k)$.

@@

Consider pseudo autocovariance and spectral variance estimators for the $s$th chain, denoted by $\tilde{\Upsilon}_s(k)$ and $\tilde{\Sigma}_s$ that use data centered around the unobserved actual mean $\mu_g$:
\begin{align*}
    \tilde{\Upsilon}_s(k) &= \dfrac{1}{n}\sum_{t \in I_k}(Y_{st}-\mu_g)(Y_{s(t+k)}-\mu_g)^T \\ 
    \tilde{\Sigma}_s &= \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\tilde{\Upsilon}_s(k) \,.
\end{align*}
%
The average pseudo spectral variance estimator is
\[
\tilde{\Sigma}_A = \dfrac{1}{m}\sum\limits_{s=1}^{m}\tilde{\Sigma}_s
\]

Further, let
\begin{align*}
  M_1 & = \dfrac{1}{m}\sum\limits_{s=1}^{m}\left\{\sum\limits_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\sum\limits_{t \in I_k}\dfrac{1}{n}\left[ \left(Y_{st}-\mu_g \right)_i   \left(\mu_g-\bar{\bar{Y}} \right)_j +    \left(\mu_g-\bar{\bar{Y}} \right)_i  \left(Y_{s(t+k)}-\mu_g \right)_j \right]\right\}\,, \\ 
  %
M_2 &= \left(\mu_g-\bar{\bar{Y}} \right)_i   \left(\mu_g-\bar{\bar{Y}} \right)_j\sum\limits_{k=-b_n+1}^{b_n-1}\left(1-\dfrac{\abs{k}}{n}\right)w\left(\dfrac{k}{b_n}\right)\,.
\end{align*}

@@

Under the assumptions of Theorem~\ref{thm:kuelbs}, $\psi(n) = o(n^{1/2})$. Using Law of Iterative Logarithms (LIL), a tighter bound for $\psi(n)$ is given by \cite{stra:1964} as $o(\sqrt{n\log \log n})$. We have already assumed that $b_n \log \log n / n \to 0$ as $n \to \infty$. Consequently, 

@@

\begin{proof}[Proof of theorem \ref{th:consistency}]

@@

% \begin{proof}[Proof of theorem \ref{th:G-SVE_bias}]
% The convoluted end effect terms in theorem \ref{th:G-_expec} are all essentially $\mathcal{O}(1/n)$. We are interested in finding the asymptotic bias here. Therefore we can write the expectation of replicated autocovariance in form of 
By Equation~\ref{eq:G-ACF_expec_breakdown},
%\begin{align*}
%\mathbb{E} \left[\hat{\Upsilon}_{G}(k) \right]
% &  = \left(1- \dfrac{k}{n}\right)\Upsilon(k) - \dfrac{k}{n}\left[\left(1-\dfrac{1}{m}\right)\left(\dfrac{1}{n}\sum_{h=0}^{n-1}\Upsilon(h)^T + \dfrac{1}{n}\sum_{h=0}^{n-1}\Upsilon(h)\right) - \left(2-\dfrac{1}{m}\right) \Var(\bar{X}_1)\right] - \dfrac{\Var(\bar{X}_1)}{m} \\
% & = \left(1 - \dfrac{\abs{k}}{n}\right)\Upsilon(k) + \mathcal{O} \left(n^{-1} \right)  \numberthis \label{eq:G-_bias2}  \,.
% \end{align*}
%

\[
\mathbb{E} \left[\hat{\Upsilon}_{G}(k) \right] = \left(1 - \dfrac{\abs{k}}{n}\right)\Upsilon(k) + O_1 + O_2    \,.
\]
%
where both $O_1 \textrm{ and } O_2$ are the small order terms where $O_1 = n^{-1}\abs{k} \mathcal{O}(n^{-1}) \text{ and } O_2 = \mathcal{O}(n^{-1})$. By our assumptions, $\sum_{k=-\infty}^{\infty}\Upsilon(k) < \infty$.  
For a truncation point $b_n$, $w(k/b_n)=0$ for all $\abs{k} > b_n$. Therefore, SV estimator can be written as a weighted sum of estimated autocovariances from lag $-n+1$ to $n-1$. Consider the G-SV estimator,
\begin{align*}
 \mathbb{E} \left[\hat{\Sigma}_{G} - \Sigma \right] & = \sum_{k=-n+1}^{n-1} w\left(\dfrac{k}{b_n}\right)\mathbb{E} \left[\hat{\Upsilon}_{G}(k) \right] - \sum_{k=-\infty}^{\infty}\Upsilon(k)\\
    &= \sum_{k=-n+1}^{n-1}  w\left(\dfrac{k}{b_n}\right)\left[\left(1-\dfrac{k}{n}\right)\Upsilon(k) + O_1 + O_2 \right]  - \sum_{k=-\infty}^{\infty}\Upsilon(k)\\
    &= \sum_{k=-n+1}^{n-1} \left[ w\left(\dfrac{k}{b_n}\right)\left(1-\dfrac{\abs{k}}{n}\right)\Upsilon(k)\right]  - \sum_{k=-\infty}^{\infty}\Upsilon(k) + \sum_{k=-n+1}^{n-1}\left[  w\left(\dfrac{k}{b_n}\right)  (O_1 + O_2) \right] \\ 
    & = P_1 + P_2\,,
\end{align*}
%
where 
\begin{align*}
    P_1 &= \sum\limits_{k=-n+1}^{n-1}\left[w\left(\dfrac{k}{b_n}\right)\left(1-\dfrac{\abs{k}}{n}\right)\Upsilon(k) \right] - \sum\limits_{k=-\infty}^{\infty}\Upsilon(k) \textrm{ and }\\   
    P_2 &= \sum_{k=-n+1}^{n-1}\left[  w\left(\dfrac{k }{b_n}\right)\left(O_1 + O_2\right)\right]\,.
\end{align*}

@@

 Due to the strong consistency proof from theorem \ref{th:consistency}, as $n \to \infty$,

@@

\section{Additional Examples}
We present two additional real-world examples to illustrate the advantage of global autocovariance estimator using ACF plots. For both the examples, we use a pre-established Markov chain structure and implement our estimator on the Markov chains obtained.
% We have established that estimating autocovariance is crucial to estimate the Markov chain standard error.  In this section we will use two real-life datsets to exemplify the better performance of our replicated autocovariance estimator for multiple Markov chains. R packages already exist to create Markov chains for these examples. However, when these chains are started from different points, we will observe that they do not represent the target distribution well for finite samples. The traditional empirical ACF estimator fails in these cases, whereas our estimator shows promising results. We will not engage into full error analysis for these examples due to computational restrictions posed by high dimensional sample space. 
\subsection{Bayesian Poisson Change Point Model}

@@

This is a 7-dimensional estimation problem wherein the marginal distribution of majority of components display a multimodal nature. Figure~\ref{fig:poisson-trace} shows the evolution of two chains started from random points with time for the second and third component. We report the ACF plots for the second component component only, however, similar behavior is observed in ACF plots of other components as well.

\begin{figure}[htbp]
    \centering
    \includegraphics[width = 0.7\textwidth]{plots/poisson-trace_n50000.pdf}
    \caption{Trace plot for second (left) and third (right) component of two chains started from random points. The two colors denote the two different chains.}
    \label{fig:poisson-trace}
\end{figure}
\begin{figure}[htbp]
    \centering
      \includegraphics[width = .6\textwidth]{plots/poisson-acf_n1000.pdf}  \\ \vspace{-.5cm}
      \includegraphics[width = .6\textwidth]{plots/poisson-acf_n10000.pdf}      
    \caption{Change point: ACF plots using local (left) and global (right) centering for $m=2$ parallel Markov chains. The individual chains are shown through dashed lines and average over $m$ chains is the solid line. The chain lengths are $n = 1000$ (top row) and $n = 10000$ (bottom row).}
    \label{fig:poisson-acf}
\end{figure}

Figure~\ref{fig:poisson-acf} demonstrates a striking advantage of G-ACF against locally-centered ACF in estimating the autocorrelations. For a chain length of $n=1000$, locally-centered ACF gives a false sense of security about the Markov chains whereas in reality, the process is highly correlated. G-ACF gives a more accurate estimation of autocorrelations in almost ten times lesser samples per chain.

@@

The \texttt{faux.magnolia.high} dataset available in the \texttt{ergm} \texttt{R} package represents a simulated within school friendship network based on Ad-Health data (\cite{resnick1997protecting}). The school communities represented by the network data are located in the southern United States. Each node represents a student and each edge represents a friendship between the nodes it connects.

@@

We believe that the students from different races experience different socio-political conditions which causes them to engage in a "selective networking". This might cause formation of clusters in the social network. We sample two parallel Markov chains starting from two students belonging to different races and study its impacts on the average features of their immediate social group, especially their grades. Due to cluster formation in the network, a Markov chain will land on students within a cluster more often than on students outside it. As a consequence, the sample averages for the two chains will differ for a relatively shorter chain length.

Figure~\ref{fig:magnolia-acf} shows the ACF plots for the \textit{grade} feature at two different simulation sizes for $m=2$ parallel Markov chains. For a shorter chain length, both the chains have explored different clusters and as a consequence, the local and global mean do not agree. Regardless, G-ACF displays a clear advantage over locally-centered ACF for a chain length of $n=100$ samples whereas the latter takes $n=1000$ samples to reach the truth.

>>>>

By \cite[Proposition 1]{song1995optimal},
\[
\Var(\bar{X}) = \dfrac{\Sigma}{n} + \dfrac{\Phi}{n^2} + o(n^{-2})\,.
\]
As a consequence, we get the result 
\[
 \E \left[\hat{\Gamma}_{G,s}(k) \right] =  \left(1- \dfrac{\abs{k}}{n}\right)\left( \Gamma(k) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2} \right) + o(n^{-1})\,.
\]
% As a consequence, if $\Var(\bar{X})$ is finite, then $\Var(\bar{X}) \to 0 \textrm{ as } n \to \infty$. Expectation of $\hat{\Gamma}_{G,s}(k)$ can then broken down as following,
Although this completes the proof of Theorem~\ref{th:G-ACF_bias}, we will require the following decomposition
 \begin{equation} \label{eq:G-ACF_expec_breakdown}
     \mathbb{E}\left[\hat{\Gamma}_{G,s}(k)\right] = \left(1- \dfrac{\abs{k}}{n}\right)\Gamma(k) + O_1 + O_2\,.
 \end{equation}
%
where,
\begin{align*}
    O_1 &= -\dfrac{\abs{k}}{n}\left[\left(1-\dfrac{1}{m}\right)\left(\dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h)^T + \dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h)\right) - \left(2-\dfrac{1}{m}\right) \left(\dfrac{\Sigma}{n} + \dfrac{\Phi}{n^2}\right)\right] + o(n^{-2})\, ,\\
    O_2 &= -\dfrac{1}{m}\left(\dfrac{\Sigma}{n} + \dfrac{\Phi}{n^2}\right) + o(n^{-2})
\end{align*}
%
% Here $O_1 = (\abs{k}/n)\mathcal{O}(1/n) \to 0 \textrm{ and } O_2 = {\mathcal{O}}(1/n) \to 0$ as $n \to \infty$. For a diagonal element of $\Gamma$,
% We observe that this expectation result is very similar to equation \ref{eq:priestly} from \cite{priestley1981spectral} for the asymptotically unbiased estimator of autocovariance except for an additional positive bias term. This helps in countering the inherent negative bias in $\hat{\Gamma}(k)$ and give more informative estimates of autocovariance.  Although, bias term for a fixed $k$ is $o(1)$, it increases with $k$. We have explicitly written the end effect term arising from using the estimator $\bar{\bar{X}}$ for $\mu$. These are small order terms that vanish as $n \to \infty$.
%
% \begin{align*}
% & \E \left[ \hat{\Gamma}_{G,s}^{ii}\right]\\
% & = \mathbb{E}\left[\hat{\Gamma}^{ii}_{s}(k)\right] - \dfrac{\abs{k}}{n}\left[\left(1-\dfrac{1}{m}\right)\left(\dfrac{1}{n}\sum_{h=0}^{n-1} \left[\Gamma^{ii}(h)^T + \Gamma^{ii}(h) \right] \right) - \left(2-\dfrac{1}{m}\right) \Var(\bar{X}_1)^{ii}\right] - \dfrac{\Var(\bar{X}_1)^{ii}}{m}\,.
% \end{align*}
% Note that $ \hat{\Gamma}_{G,s}(k)$ is less negatively biased as compared to $\hat{\Gamma}_s(k)$. In the presence of positive correlation, the leftover $o(n^{-1})$ term is positive which further counters the negative bias of $\hat{\Gamma}_{G,s}(k)$.

@@

Consider pseudo autocovariance and spectral variance estimators for the $s$th chain, denoted by $\tilde{\Upsilon}_s(k)$ and $\tilde{\Sigma}_s$ that use data centered around the unobserved actual mean $\mu_g$:
\begin{align*}
    \tilde{\Upsilon}_s(k) &= \dfrac{1}{n}\sum_{t \in I_k}(Y_{st}-\mu_g)(Y_{s(t+k)}-\mu_g)^T  \quad \text{and} \quad 
    \tilde{\Sigma}_s = \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\tilde{\Upsilon}_s(k) \,.
\end{align*}
%
The average pseudo spectral variance estimator is $\tilde{\Sigma}_A = m^{-1}\sum_{s=1}^{m}\tilde{\Sigma}_s$
% \[
% \tilde{\Sigma}_A = \dfrac{1}{m}\sum\limits_{s=1}^{m}\tilde{\Sigma}_s
% \]
% 
Further, let
\begin{align*}
  M_1 & = \dfrac{1}{m}\sum\limits_{s=1}^{m}\left\{\sum\limits_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\sum\limits_{t \in I_k}\dfrac{1}{n}\left[ \left(Y_{st}-\mu_g \right)_i   \left(\mu_g-\bar{\bar{Y}} \right)_j +    \left(\mu_g-\bar{\bar{Y}} \right)_i  \left(Y_{s(t+k)}-\mu_g \right)_j \right]\right\}\,, \\ 
  %
M_2 &= \left(\mu_g-\bar{\bar{Y}} \right)_i   \left(\mu_g-\bar{\bar{Y}} \right)_j\sum\limits_{k=-b_n+1}^{b_n-1}\left(1-\dfrac{\abs{k}}{n}\right)w\left(\dfrac{k}{b_n}\right)\,.
\end{align*}

@@

Under the assumptions of Theorem~\ref{thm:kuelbs}, $\psi(n) = o(n^{1/2})$. Using the law of iterative logarithms (LIL), a tighter bound for $\psi(n)$ is given by \cite{stra:1964} as $o(\sqrt{n\log \log n})$. Since $b_n \log \log n / n \to 0$ as $n \to \infty$, consequently, 

@@

\begin{proof}[Proof of Theorem \ref{th:consistency}]

@@

% \begin{proof}[Proof of theorem \ref{th:G-SVE_bias}]
% The convoluted end effect terms in theorem \ref{th:G-_expec} are all essentially $\mathcal{O}(1/n)$. We are interested in finding the asymptotic bias here. Therefore we can write the expectation of replicated autocovariance in form of 
By Equation~\ref{eq:G-ACF_expec_breakdown},
%\begin{align*}
%\mathbb{E} \left[\hat{\Upsilon}_{G}(k) \right]
% &  = \left(1- \dfrac{k}{n}\right)\Upsilon(k) - \dfrac{k}{n}\left[\left(1-\dfrac{1}{m}\right)\left(\dfrac{1}{n}\sum_{h=0}^{n-1}\Upsilon(h)^T + \dfrac{1}{n}\sum_{h=0}^{n-1}\Upsilon(h)\right) - \left(2-\dfrac{1}{m}\right) \Var(\bar{X}_1)\right] - \dfrac{\Var(\bar{X}_1)}{m} \\
% & = \left(1 - \dfrac{\abs{k}}{n}\right)\Upsilon(k) + \mathcal{O} \left(n^{-1} \right)  \numberthis \label{eq:G-_bias2}  \,.
% \end{align*}
%
%
\[
\mathbb{E} \left[\hat{\Upsilon}_{G}(k) \right] = \left(1 - \dfrac{\abs{k}}{n}\right)\Upsilon(k) + O_1 + O_2    \,.
\]
%
where both $O_1 \textrm{ and } O_2$ are the small order terms where $O_1 = n^{-1}\abs{k} \mathcal{O}(n^{-1}) \text{ and } O_2 = \mathcal{O}(n^{-1})$. By our assumptions, $\sum_{k=-\infty}^{\infty}\Upsilon(k) < \infty$.  
For a truncation point $b_n$, $w(k/b_n)=0$ for all $\abs{k} > b_n$. Therefore, SV estimator can be written as a weighted sum of estimated autocovariances from lag $-n+1$ to $n-1$. Consider the G-SV estimator,
\begin{align*}
 \mathbb{E} \left[\hat{\Sigma}_{G} - \Sigma \right] & = \sum_{k=-n+1}^{n-1} w\left(\dfrac{k}{b_n}\right)\mathbb{E} \left[\hat{\Upsilon}_{G}(k) \right] - \sum_{k=-\infty}^{\infty}\Upsilon(k)\\
    &= \sum_{k=-n+1}^{n-1}  w\left(\dfrac{k}{b_n}\right)\left[\left(1-\dfrac{k}{n}\right)\Upsilon(k) + O_1 + O_2 \right]  - \sum_{k=-\infty}^{\infty}\Upsilon(k)\\
    &= \sum_{k=-n+1}^{n-1} \left[ w\left(\dfrac{k}{b_n}\right)\left(1-\dfrac{\abs{k}}{n}\right)\Upsilon(k)\right]  - \sum_{k=-\infty}^{\infty}\Upsilon(k) + \sum_{k=-n+1}^{n-1}\left[  w\left(\dfrac{k}{b_n}\right)  (O_1 + O_2) \right] \\ 
    & = P_1 + P_2\,,
\end{align*}
%
where 
\begin{align*}
    P_1 &= \sum\limits_{k=-n+1}^{n-1}\left[w\left(\dfrac{k}{b_n}\right)\left(1-\dfrac{\abs{k}}{n}\right)\Upsilon(k) \right] - \sum\limits_{k=-\infty}^{\infty}\Upsilon(k) \textrm{ and }\\   
    P_2 &= \sum_{k=-n+1}^{n-1}\left[  w\left(\dfrac{k }{b_n}\right)\left(O_1 + O_2\right)\right]\,.
\end{align*}

@@

 Due to the strong consistency proof from Theorem \ref{th:consistency}, as $n \to \infty$,

@@

\section{Additional Examples}
We present two additional examples to illustrate the advantage of the G-ACF estimator. 
% For both the examples, we use a pre-established Markov chain structure and implement our estimator on the Markov chains obtained.
% We have established that estimating autocovariance is crucial to estimate the Markov chain standard error.  In this section we will use two real-life datsets to exemplify the better performance of our replicated autocovariance estimator for multiple Markov chains. R packages already exist to create Markov chains for these examples. However, when these chains are started from different points, we will observe that they do not represent the target distribution well for finite samples. The traditional empirical ACF estimator fails in these cases, whereas our estimator shows promising results. We will not engage into full error analysis for these examples due to computational restrictions posed by high dimensional sample space. 
\subsection{Bayesian Poisson Change Point Model}

@@

This yields a 7-dimensional posterior distribution in $\lambda = (\lambda_1, \dots, \lambda_7)^T$. Figure~\ref{fig:poisson-trace} shows the evolution of two chains started from random points for the second and third component. We report the ACF plots for the second component component only, however, similar behavior is observed in ACF plots of other components as well.

\begin{figure}[htbp]
    \centering
    \includegraphics[width = 0.7\textwidth]{plots/poisson-trace_n50000.pdf}
    \caption{Trace plot for second (left) and third (right) component of two chains.}
    \label{fig:poisson-trace}
\end{figure}
\begin{figure}[htbp]
    \centering
      \includegraphics[width = .6\textwidth]{plots/poisson-acf_n1000.pdf}  \\ \vspace{-.5cm}
      \includegraphics[width = .6\textwidth]{plots/poisson-acf_n10000.pdf}      
    \caption{Change point: ACF plots using local (left) and global (right) centering for $m=2$ parallel Markov chains. The individual chains are shown through dashed lines and average over $m$ chains is the solid line. The chain lengths are $n = 1000$ (top row) and $n = 10000$ (bottom row).}
    \label{fig:poisson-acf}
\end{figure}

Figure~\ref{fig:poisson-acf} demonstrates a striking advantage of G-ACF against locally-centered ACF in estimating the autocorrelations. For a chain length of $n=1000$, locally-centered ACF gives a false sense of security about the Markov chains whereas in reality, the process is highly correlated. G-ACF gives a more accurate estimation of autocorrelations even at short run lengths.

@@

The \texttt{faux.magnolia.high} dataset available in the \texttt{ergm} \texttt{R} package represents a simulated friendship network based on Ad-Health data (\cite{resnick1997protecting}). The school communities represented by the network data are located in the southern United States. Each node represents a student and each edge represents a friendship between the nodes it connects.

@@

% We believe that the students from different races experience different socio-political conditions which causes them to engage in a "selective networking". This might cause formation of clusters in the social network. 
We sample two parallel Markov chains starting from two students belonging to different races.
 % and study its impacts on the average features of their immediate social group, especially their grades. Due to cluster formation in the network, a Markov chain will land on students within a cluster more often than on students outside it. As a consequence, the sample averages for the two chains will differ for a relatively shorter chain length. 
% 
Figure~\ref{fig:magnolia-acf} shows the ACF plots for the \textit{grade} feature at two different simulation sizes for $m=2$ parallel Markov chains. For a shorter chain length, both the chains have explored different clusters and as a consequence, the local and global mean do not agree. Regardless, G-ACF displays a clear advantage over locally-centered ACF for a chain length of $n=100$ samples whereas the latter takes $n=1000$ samples to reach the truth.
